Nonlinear methods to quantify Movement Variability in Human-Humanoid Interaction Activities

### 2021-04-05 docs/manuscript/main.tex
#13 addresses reference of figures; and adds workflows; README details and sh for local build

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\subsection*{Structures of Recurrence Plots}
Pattern formations in the RPs can be designated either as topology for large-scale patterns or texture for small-scale patterns.
In the case of topology, the following pattern formations are commonly presented: (i) homogeneous where uniform recurrence points are spread in the RP e.g., uniformly distributed noise (Fig~\ref{fig01}c3), (ii) periodic and quasi-periodic systems where diagonal lines and checkerboard structures represent oscillating systems, e.g., sinusoidal signals (Fig~\ref{fig01}c4), (iii) drift where paling or darkening recurrence points away from the LOI is caused by drifting systems, e.g., logistic map (Fig~\ref{fig01}c5), and (iv) disrupted where recurrence points are presented white areas or bands that indicate abrupt changes in the dynamics, e.g. Brownian motion (Fig~\ref{fig01}c7) \cite{eckmann1987, marwan2015}.
Texture patterns in RPs can be categorised as: (i) single or isolated recurrence points that represent rare occurring states, and do not persist for any time or fluctuate heavily, (ii) dots forming diagonal lines where the length of the small-scale parallel lines in the diagonal are related to the ratio of determinism or predictability in the dynamics of the system, and (iii) dots forming vertical and horizontal lines where the length of the lines represent a time length where a state does not change or change very slowly and these patterns formation represent discontinuities in the signal, and (iv) dots clustering to inscribe rectangular regions which are related to laminar states or singularities \cite{marwan2015}.

@@

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\section*{Experiment} \label{sec:experiment}
We conducted an experiment in the context of human-humanoid imitation (HHI) activities where participants were asked to imitate simple horizontal and vertical arm movements performed by NAO, a humanoid robot \cite{gouaillier2009}.
Such simple movements were repeated ten times for the participant who copied NAO's arm movements in a face-to-face imitation activity.
Also, wearable inertial measurement unit (IMU) sensors were attached to the right hand of the participant and to the left hand of the robot (Figure~\ref{fig02}a1,a3).
Data were then collected with four NeMEMSi IMU sensors with sampling rate of 50Hz provinding tri-axial data of the accelerometer, gyroscope and magnetometer sensors and quaternions \cite{Comotti2014}.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
  \centering
\includegraphics[width=1.0\textwidth]{figures/fig02/versions/drawing-v01.png}
    \caption{
	{\bf Human-humanoid imitation activities (A). Time series duration for arm movements (B).
		Time series for horizontal and vertical arm movements (C).
	}
	Face-to-face human-humanoid imitation (HHI) activities for (a1) HHI of horizontal arm movement, (a2) Humanoid horizontal arm movement, (a3) HHI of vertical arm movement, and (a4) Humanoid vertical arm movement.
    Time series of smoothed data from gyroscope sensor for different speed arm movements performed by NAO: (b1) Horizontal Normal arm movement (HN), (b2) Horizontal Faster arm movement (HF), (b3) Vertical Normal arm movement (VN) and (b4) Vertical Faster arm movement (VF).
	(B) shows window sizes for 2-seconds (100 samples), 5-seconds (250 samples), 10-seconds (500 samples) and 15-seconds (750 samples).
	(c1,c4) raw-normalised (sg0zmuv), (c2,c5) normalised-smoothed 1 (sg1zmuv) and (c3,c7) normalised-smoothed 2 (sg2zmuv).
	Time series are only for three participants	($p01$, $p02$, and $p03$) for horizontal and vertical arm movements in normal and faster velocity (HN, HF, VN, VF) with the normalised GyroZ or GyroY axis and with one sensor attached to the participant (HS01) and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2021}.
}
    \label{fig02}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

@@

\subsection*{Human-humanoid imitation activities}
For human-humanoid imitation (HHI) activities four neMEMSi sensors were used, two of which were attached to the right hand of the participant and the other two to the left hand of the humanoid robot.
Then, each participant was asked to imitate repetitions of simple horizontal and vertical arm movements performed by the humanoid robot in the following conditions:
(i) ten repetitions of horizontal arm movement at normal (HN) and faster (HF) speed (Figure~\ref{fig02}a1), and
(ii) ten repetitions of vertical arm movement at normal (VN) and faster (VF) speed (Figure~\ref{fig02}a3).
The normal and faster speed of arm movements is defined by the duration in number of samples of one repetition of NAO's arm movements.
We select NAO's arm movements duration to distinguish between normal and faster arm movements as NAO's movements have less variation for such repetitive movements.
The duration for one repetition of the horizontal arm movement at normal speed, HN, is 5 seconds considering that each repetition last around 250 samples.
For horizontal arm movement at faster speed, HF, each repetition were performed in 2 seconds which correspond to 90 samples of data.
The vertical arm movement at normal speed, VN, were performed  in 6 seconds which is around 300 samples of data.
For vertical arm movement at faster speed, VF, each repetition lasts about 2.4 seconds which correspond to 120 samples of data.
To visualise the distinction between normal and faster speed for horizontal and vertical arm movements, Fig~\ref{fig02}B shows smoothed time series for axes Z and Y of the gyroscope sensors with four window lengths: 2-sec (100-samples), 5-sec (250-samples), 10-sec (500-samples) and 15-sec (750-samples).

>>>>

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\subsection*{Structures of Recurrence Plots}
Pattern formations in the RPs can be designated either as topology for large-scale patterns or texture for small-scale patterns.
In the case of topology, the following pattern formations are commonly presented: (i) homogeneous where uniform recurrence points are spread in the RP e.g., uniformly distributed noise (Fig~\ref{fig01}c3), (ii) periodic and quasi-periodic systems where diagonal lines and checkerboard structures represent oscillating systems, e.g., sinusoidal signals (Fig~\ref{fig01}c4), (iii) drift where paling or darkening recurrence points away from the LOI is caused by drifting systems, e.g., logistic map (Fig~\ref{fig01}c5), and (iv) disrupted where recurrence points are presented white areas or bands that indicate abrupt changes in the dynamics, e.g. Brownian motion (Fig~\ref{fig01}c6) \cite{eckmann1987, marwan2015}.
Texture patterns in RPs can be categorised as: (i) single or isolated recurrence points that represent rare occurring states, and do not persist for any time or fluctuate heavily, (ii) dots forming diagonal lines where the length of the small-scale parallel lines in the diagonal are related to the ratio of determinism or predictability in the dynamics of the system, and (iii) dots forming vertical and horizontal lines where the length of the lines represent a time length where a state does not change or change very slowly and these patterns formation represent discontinuities in the signal, and (iv) dots clustering to inscribe rectangular regions which are related to laminar states or singularities \cite{marwan2015}.

@@

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\section*{Experiment} \label{sec:experiment}
We conducted an experiment in the context of human-humanoid imitation (HHI) activities where participants were asked to imitate simple horizontal (Figure~\ref{fig02}a1) and vertical arm movements (Figure~\ref{fig02}a3) performed by NAO, a humanoid robot (Figure~\ref{fig02}a2,~\ref{fig02}a4) \cite{gouaillier2009}.
Such simple movements were repeated ten times for the participant who copied NAO's arm movements in a face-to-face imitation activity.
Also, wearable inertial measurement unit (IMU) sensors were attached to the right hand of the participant and to the left hand of the robot (Figures~\ref{fig02}a1,~\ref{fig02}a3).
Data were then collected with four NeMEMSi IMU sensors with sampling rate of 50Hz provinding tri-axial data of the accelerometer, gyroscope and magnetometer sensors and quaternions \cite{Comotti2014}.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
  \centering
\includegraphics[width=1.0\textwidth]{figures/fig02/versions/drawing-v01.png}
    \caption{
	{\bf Human-humanoid imitation activities (A). Time series duration for arm movements (B).
		Time series for horizontal and vertical arm movements (C).
	}
	Face-to-face human-humanoid imitation (HHI) activities for (a1) HHI of horizontal arm movement, (a2) Humanoid horizontal arm movement, (a3) HHI of vertical arm movement, and (a4) Humanoid vertical arm movement.
    Time series of smoothed data from gyroscope sensor for different speed arm movements performed by NAO: (b1) Horizontal Normal arm movement (HN), (b2) Horizontal Faster arm movement (HF), (b3) Vertical Normal arm movement (VN) and (b4) Vertical Faster arm movement (VF).
	(B) shows window sizes for 2-seconds (100 samples), 5-seconds (250 samples), 10-seconds (500 samples) and 15-seconds (750 samples).
	(c1,c4) raw-normalised (sg0zmuv), (c2,c5) normalised-smoothed 1 (sg1zmuv) and (c3,c7) normalised-smoothed 2 (sg2zmuv).
	Time series are only for three participants	($p01$, $p02$, and $p03$) for horizontal and vertical arm movements in normal and faster velocity (HN, HF, VN, VF) with the normalised GyroZ or GyroY axis and with one sensor attached to the participant (HS01) and other sensor attached to the robot (RS01).
	Code and data to reproduce the figure is available in \cite{srep2021}.
}
    \label{fig02}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

@@

\subsection*{Human-humanoid imitation activities}
For human-humanoid imitation (HHI) activities four neMEMSi sensors were used, two of which were attached to the right hand of the participant and the other two to the left hand of the humanoid robot (Figure~\ref{fig02}A).
Then, each participant was asked to imitate repetitions of simple horizontal and vertical arm movements performed by the humanoid robot in the following conditions:
(i) ten repetitions of horizontal arm movement at normal (HN) and faster (HF) speed (Figure~\ref{fig02}a1), and
(ii) ten repetitions of vertical arm movement at normal (VN) and faster (VF) speed (Figure~\ref{fig02}a3).
The normal and faster speed of arm movements is defined by the duration in number of samples of one repetition of NAO's arm movements.
We select NAO's arm movements duration to distinguish between normal and faster arm movements as NAO's movements have less variation for such repetitive movements.
The duration for one repetition of the horizontal arm movement at normal speed, HN, is 5 seconds considering that each repetition last around 250 samples.
For horizontal arm movement at faster speed, HF, each repetition were performed in 2 seconds which correspond to 90 samples of data.
The vertical arm movement at normal speed, VN, were performed  in 6 seconds which is around 300 samples of data.
For vertical arm movement at faster speed, VF, each repetition lasts about 2.4 seconds which correspond to 120 samples of data.
To visualise the distinction between normal and faster speed for horizontal and vertical arm movements, Fig~\ref{fig02}B shows smoothed time series for axes Z and Y of the gyroscope sensors with four window lengths: 2-sec (100-samples), 5-sec (250-samples), 10-sec (500-samples) and 15-sec (750-samples).
